Lattices of acyclic pipe dreams

### 2023-03-08 acyclicFacets/draft/latticePipeDreams.tex
sweeping algorithm

We refer to~\cite{BjornerBrenti, Humphreys} for detailed references on Coxeter groups. 
We consider a \defn{finite root system} $\Phi$ with \defn{positive roots}~$\Phi^+$, \defn{negative roots}~$\Phi^+-$, and \defn{simple roots} $\Delta \subseteq \Phi^+$.
The reflections along the hyperplanes orthogonal to the roots in $\Phi$ generate a \defn{finite Coxeter group} $W$.
We denote by~$s_\alpha \in W$ the reflection orthogonal to a root~$\alpha \in \Phi$, and by~$\alpha_s \in \Phi^+$ the positive root orthogonal to a reflection~$s \in W$.
The group~$W$ is actually generated by the simple reflections~$S = \set{s_\alpha}{\alpha \in \Delta}$, and the pair~$(W,S)$ is a \defn{finite Coxeter system}.
Finite root systems and finite Coxeter systems are classified in terms of Dynkin diagrams, see~\cite{Humphreys}.

@@

It is known that~$\subwordComplex(Q,\omega)$ is either a ball or a sphere, in particular it is a pseudomanifold (with or without boundary).
The flip graph of~$\subwordComplex(Q,\omega)$ is the graph whose vertices are the facets of~$\subwordComplex(Q,\omega)$ and whose edges are the ridges of~$\subwordComplex(Q,\omega)$.
In other words, two facets $I,J \in \subwordFacets(Q,\omega)$ are connected by a \defn{flip} if they differ one element: $I \ssm \{i\} = J \ssm \{j\}$ for some $i\in I$ and $j\in J$ with~$i \ne j$.
The flip from $I$ to $J$ is called \defn{increasing} if~$i < j$, and \defn{decreasing} otherwise.
The \defn{increasing flip graph} on the facets of the subword complex is an acyclic graph which has a unique source and a unique sink~\cite{Pilaud-greedyFlipTree, PilaudStump-ELlabelings}.
These two special facets are called the \defn{greedy facet} $\greedyFacet$ and the \defn{antigreedy facet} $\antiGreedyFacet$.
They are the lexicographically smallest and largest facets of $\subwordComplex(Q,\pi)$, respectively.
The \defn{increasing flip poset} is the transitive closure of the increasing flip graph.

@@

\begin{example}
\label{exm:typeARootConfiguration}
Continuing \cref{exm:typeACoxeterSystem}, we consider the type~$A_n$ Coxeter system, the word $Q \eqdef \tau_1 \cdots \tau_{n-1} \tau_1 \cdots \tau_{n-2} \cdots \tau_1 \tau_2 \tau_1$, and a permutation~$\omega \in \fS_n$.
The word~$Q$ naturally fits on an $n \times n$ triangular grid (place $\tau_k$ in all boxes with row~$i$ and column~$j$ such that~$i+j+k = n+1$) and each facet~$I$ of the subword complex~$\subwordComplex(Q, \omega)$ corresponds to a pipe dreams~$P_I$ of~$\pipeDreams(\omega)$ (replace each position in~$I$ by a contact in~$P_I$, and the other positions by crossings in~$P_I$).
The root function is given~$\rootFunction{I}{i} = \b{r}_{p,q} \eqdef \b{e}_p - \b{e}_q$ where~$p$ is the pipe arriving from the west and $q$ is the pipe arriving from the south at the box of~$P_I$ corresponding to position~$i$ of~$Q$.
Hence, the root configuration is the incidence configuration~$\Roots(I) = \bigset{\b{r}_{p,q}}{(p,q) \in P_I\contact}$ of the contact graph of~$P_I$.
In particular, the acyclic facets of~$\subwordComplex(Q, \omega)$ correspond to the acyclic pipe dreams of~$\acyclicPipeDreams(\omega)$.
\end{example}

@@

\begin{example}
\label{exm:counterExampleLattice}
In contrast to the previous example, consider the word~$Q = \tau_1 \tau_2 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_2 \tau_1$ on the simple generators of the symmetric group~$\fS_4$ and the subword complex~$\subwordComplex(Q,\wo)$.
Note that all facets in this subword complex are acyclic.
It is shown in~\cite[Rem.~5.12]{PilaudStump-brickPolytope} that its increasing flip poset is not a lattice.
This shows that neither the increasing flip poset, nor the acyclic increasing flip poset are lattices in general, even in type~$A$. \vincent{clean that}
\end{example}

@@

\begin{proof}
%Let $\Phi = \Phi^+ \sqcup \Phi^-$ be the root system associated to $W$.
A classical result for finite root systems states that for any generic linear halfspace~$H^+$ the intersection~$\Phi \cap H^+$ is of the form $\pi(\Phi^+)$ for some $\pi \in W$.
Since any pointed cone is contained in some generic linear halfspace, we obtain
\[
\begin{array}{rcl}
I \text{ is acyclic}  & \longleftrightarrow  &  \cone \Roots(I) \text{ is pointed} \\
  & \longleftrightarrow  & \cone \Roots(I)\subseteq \pi(\Phi^+) \text{ for some } \pi\in W \\
  & \longleftrightarrow  &  \Roots(I)\subseteq \pi(\Phi^+) \text{ for some } \pi\in W \\
  & \longleftrightarrow  &  \linearExtensions(I) \neq \varnothing 
  \qedhere
\end{array}
\]
\end{proof}

@@

A subset $\fR$ of the weak order is 
\begin{itemize}
\item a \defn{lower set} if $\sigma < \tau$ and $\tau \in \fR$ implies~$\sigma \in \fR$,
\item \defn{order convex} if $\sigma < \tau < \rho$ and~$\sigma, \rho \in \fR$ implies~$\tau \in \R$.
\end{itemize}

@@

Now, since $\beta\notin \pi(\Phi^+)$.
There must be only one $\beta$ in the list.
Otherwise, there would be at least one $\beta$ whose position belongs to the facet $I$.
This would imply that $\beta\in \Roots(I)$ and $\pi$ would not be a linear extension of $I$, which is a contradiction.
So, our restricted list corresponding to $I$ looks like
\begin{equation}
\begin{array}{ccc}
  \beta &, -\beta, \dots, &-\beta.\\
      i && j
\end{array}    
\end{equation}
By  \cref{lem:rootFunctionFlips}~\eqref{lem:rootFunctionFlips3}, we obtain that flipping $j$ to $i$ creates the new facet $I' = I \ssm \{j\} \cap \{i\}$ whose corresponding restricted list looks like
\begin{equation}
\begin{array}{ccc}
  \beta &, \beta, \dots, &\beta.\\
      i && j
\end{array}    
\end{equation}

@@

In this section, we present conjectural generalizations of the results of \cref{subsec:pipeDreamCongruence,subsec:pipeDreamQuotient}.
By \cref{coro:linearExtensionsPartitionSubwordComplexes}, we have
\[
\linearExtensions(Q,\omega) = \bigsqcup_{I \in \subwordAcyclicFacets(Q, \omega)} \linearExtensions(I)
\]
which naturally defines an equivalence relation on~$\linearExtensions(Q,\omega)$.
However, this equivalence relation is in general not a lattice congruence for two obvious reasons:
\begin{enumerate}[(i)]
\item while~$\linearExtensions(Q,\omega)$ is a lower set of the weak order containing~$[e, \omega]$ by \cref{thm:linearExtensionsPartitionSubwordComplexA}, it does not always coincides with~$[e, \omega]$ by \cref{rem:linearExtensionsPartitionSubwordComplexA}, and in fact it does not necessarily have a maximal element, hence it is not necessarily a lattice,
\item while the sets~$\linearExtensions(I)$ are always order convex in the weak order by \cref{thm:linearExtensionsPartitionSubwordComplexA}~\eqref{item:convex}, they are not necessarily intervals in the weak order by \cref{rem:linearExtensionsPartitionSubwordComplexA}.
\end{enumerate}
To bypass Issue~(i), we could restrict our attention to the situation when the word~$Q$ is sorting by~\cref{thm:linearExtensionsPartitionSubwordComplexB} (we will do that in \cref{conj:latticeQuotientSortingAlternating}).
Here, we want to be slightly more general, so we will instead consider general words~$Q$, but restrict our attention to the interval~$[e, \omega]$ as follows.
For a facet~$I$ of~$\subwordComplex(Q, \omega)$, we define~$\strongLinearExtensions(I) \eqdef \linearExtensions(I) \cap [e, \omega]$.
We say that~$I$ is \defn{strongly acyclic}~$\strongLinearExtensions(I) \ne \varnothing$, and we denote by~$\subwordStronglyAcyclicFacets(Q, \omega)$ the set of strongly acyclic facets of~$\subwordComplex(Q, \omega)$.
Note that by \cref{thm:linearExtensionsPartitionSubwordComplexA}~\eqref{item:partition}, we have
\[
[e, \omega] = \bigsqcup_{I\in\subwordStronglyAcyclicFacets(Q,\omega)} \strongLinearExtensions(I).
\]
We can thus now define the analogue of the pipe dream congruence of \cref{def:pipeDreamCongruence} for subword complexes as follows.

@@

\subsection{The sweeping algorithm}
\label{subsec:sweepingAlgorithmSubwordComplexes}
\vincent{I have not checked this part yet.}
\vincent{Here, we must properly cite \cite{JahnStump} and say that we obtained this algorithm independently.}

Given a linear extension $\pi \in \linearExtensions(Q,\omega)$, one natural question is whether we can recover the unique acyclic facet $I$ such that $\pi \in \linearExtensions(I)$.
We can recover this facet via a sweeping algorithm which we now introduce.
The algorithm outcomes a facet for every element~$\pi \in W$, and this will be the desired facet for each $\pi \in \linearExtensions(Q,\omega)$.

\[
\begin{tabular}{cccc}
$\sweepingAlgorithm$ : & $W$  &$\longrightarrow$ &$\subwordFacets(Q,\omega)$\\
    & $\pi$ & $\longrightarrow$ & $\sweepingAlgorithm(Q,\omega,\pi)$
\end{tabular}
\]

We start by setting $I^0=\varnothing$ to be the empty set.
Then we scan the word $Q=(q_1,\dots ,q_m)$ from left to right.
At position $j\in [m]$ we produce a new set $I^j$ obtained from the previous one $I^{j-1}$ by either adding $j$ or not, according to certain rules.
We let $Q^j=Q_{[j]\ssm I^j}$ and denote by
    \[
    \rootFunction{I^j}{k} = \prod Q_{[k-1]\ssm I^j}(\alpha_{q_k}).
    \]
the partial root function for $k\leq j$.  
The rules are the following:

\begin{enumerate}
    \item If $\rootFunction{I^j}{j}\in \Ninv(\omega)$ then let 
    \[
    I^j=I^{j-1}\cup\{j\}.
    \] 
    \label{sweeping_case1}
    \item If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Inv (\pi)$ then let 
    \[
    I^j=I^{j-1}.
    \]
    \label{sweeping_case2}
    \item If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Ninv (\pi)$ then we consider two cases: 
    \label{sweeping_case3} 
    \begin{enumerate}
        \item If $Q_{[m]\ssm I^{j-1}\ssm \{j\}}$ contains a reduced expression of $\omega$ with prefix $Q^{j-1}$ then let 
        \[
        I^j=I^{j-1}\cup \{j\}.
        \]        
        \label{sweeping_case3a}
        \item If not let 
        \[
        I^j=I^{j-1}.
        \]
        \label{sweeping_case3b}
    \end{enumerate}
    \item If $\rootFunction{I^j}{j}\in -\Inv(\omega)$ then let
    \[
    I^j=I^{j-1}\cup \{j\}.
    \]
    \label{sweeping_case4}
\end{enumerate}

We denote by \defn{$\sweepingAlgorithm(Q,\omega,\pi)$} the resulting set $I^m$ obtained at the last step of the algorithm.

Roughly speaking, the goal of the algorithm is to insert a reduced expression of $\omega$ in $Q$ by sweeping the word from left to right, while deciding whether we use a letter or not by looking at the root $\beta=\rootFunction{I^j}{j}$ that it produces.
If $\beta \in \Ninv(\omega)$ then it can not be used (Case~\eqref{sweeping_case1}), otherwise we would not have a reduced expression for $\omega$.
If $\beta \in \Inv(\omega)$ then we take it when $\beta \in \Inv(\pi)$ (Case~\eqref{sweeping_case2}), or delay it as much as possible when $\beta \in \Ninv(\pi)$ (Case~\eqref{sweeping_case3}).
If $\beta \in -\Inv(\omega)$ we can simply not take it (Case~\eqref{sweeping_case4}).

\begin{proposition}
\label{prop_sweeping1}
    Let $\subwordComplex(Q,\omega)$ be a non-empty subword complex.
    For every $\pi \in W$, the set $\sweepingAlgorithm(Q,\omega,\pi)$ is a facet of $\subwordComplex(Q,\omega)$.
\end{proposition}

\begin{proof}
    Let $I^j$ and $Q^j=Q_{[j]\ssm I^j}$ be as defined in the sweeping algorithm for $\pi$. 
    The proposition follows by applying the following claim for $j=m$.
    \smallskip

    {\bf Claim}: $Q^j$ is a reduced expression which is the prefix of a reduced expression of $\omega$ in $Q_{[m]\ssm I^j}$.

    \smallskip
    We prove the claim by induction on $j$.

@@

    Now assume that the claim holds for $j-1$ and prove it for $j$.
    Note that 
    \begin{equation}
         Q^j = 
  \begin{cases}
    Q^{j-1}, & \text{if } I^j=I^{j-1}\cup \{j\} \\
    Q^{j-1}\circ (q_j), & \text{if } I^j=I^{j-1}
  \end{cases}
    \end{equation}    
    We analyze the different cases of the sweeping algorithm.
    \begin{enumerate}
    \item If $\rootFunction{I^j}{j}\in \Ninv(\omega)$ then 
    $Q^j=Q^{j-1}$, which is reduced by induction hypothesis.
    Moreover, it is a prefix of a reduced expression of $\omega$ in $Q_{[m]\ssm I^{j-1}}$.
    Since $\rootFunction{I^j}{j}\in \Ninv(\omega)$, this reduced expression can not use the letter $q_j$; so, it is a reduced expression of $\omega$ in $Q_{[m]\ssm I^{j}}$ as desired.
    \item If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Inv (\pi)$ then $Q^j= Q^{j-1}\circ (q_j)$, which is a reduced expression because~$Q^{j-1}$ is reduced and $\rootFunction{I^j}{j}\in \Inv(\omega)$.

@@

    \item[(3)(a)] If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Ninv (\pi)$ and $Q_{[m]\ssm I^{j-1}\ssm \{j\}}$ contains a reduced expression of $\omega$ with prefix $Q^{j-1}$, then $Q^j=Q^{j-1}$.
    Therefore, $Q^j$ is reduced and it is a prefix of a reduced expression of $\omega$ in 
    $Q_{[m]\ssm I^{j-1}\ssm \{j\}}=Q_{[m]\ssm I^j}$.

    \item[(3)(b)] In this case, $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Ninv (\pi)$ but $Q^j= Q^{j-1}\circ (q_j)$. The proof is similar to the proof of (2).

    \item[(4)] If $\rootFunction{I^j}{j}\in -\Inv(\omega)$ then $Q^j=Q^{j-1}$. The argument is similar to that of part (1).
    \end{enumerate}

Finally, we need to argue that any position $j$ falls into one of these cases.
The root $\rootFunction{I^j}{j}\in \Phi^+ \sqcup \Phi^-$ and we know that
\[
\Phi^+ = \Inv(\omega) \sqcup \Ninv(\omega) \qquad \qquad
\Phi^- = -\Inv(\omega) \sqcup -\Ninv(\omega)  
\]
The positive roots are covered by the cases (1),(2), and (3).
The negative roots in $-\Inv(\omega)$ are covered in case (4).
Since we are inserting a reduced expression of $\omega$ in $Q$, the case $-\Ninv(\omega)$ never occurs. 
\end{proof}

The following proposition shows that for every linear extension $\pi \in \linearExtensions(Q,\omega)$, the sweeping algorithm recovers the unique acyclic facet $I$ such that $\pi \in \linearExtensions(I)$.

\begin{proposition}
\label{prop_sweeping2}
    If $\pi\in \linearExtensions(I)$ for some facet $I\in \subwordComplex(Q,\omega)$, then 
    \[
I = \sweepingAlgorithm(Q,\omega,\pi).
    \]
\end{proposition}

Our proof is based on the following lemma.

\begin{lemma}\label{lem_negativeroot}
    Let $I_1,I_2\in \subwordComplex(Q,\omega)$ be two different facets, and $j\in [m]$ be the first position where they differ.
    Without loss of generality assume 
    \[
I_2\cap [j] = I_1\cap [j] \ssm \{j\}
    \]
with $j\in I_1$.
Let $\beta=\rootFunction{I_1}{j}=\rootFunction{I_2}{j}$, then 
\[
-\beta \in \cone \Roots(I_2).
\]
\end{lemma}

@@

\begin{proof}[Proof of \cref{prop_sweeping2}]
Assume $\pi \in \linearExtensions(I)$ and let $I^j=I\cap [j]$.
We will show that the partial root function $\rootFunction{I^j}{\cdot}$ agrees with the decisions taken in the sweeping algorithm.
Indeed, we will see that those decisions are forced.

Recall that

\begin{center}
\begin{tabular}{ccl}
    $\pi \in \linearExtensions(I)$ & $\longleftrightarrow$ & $R(I)\subseteq \pi(\Phi^+)$\\
     & $\longleftrightarrow$ & $\cone R(I)\subseteq \pi(\Phi^+)$
\end{tabular}    
\end{center}
and 
\[
\pi(\Phi^+) = -\Inv(\pi) \sqcup \Ninv(\pi).
\]
We analyze the possible cases in the sweeping algorithm.

>>>>

We refer to~\cite{BjornerBrenti, Humphreys} for detailed references on Coxeter groups. 
We consider a \defn{finite root system} $\Phi$ with \defn{positive roots}~$\Phi^+$, \defn{negative roots}~$\Phi^-$, and \defn{simple roots} $\Delta \subseteq \Phi^+$.
The reflections along the hyperplanes orthogonal to the roots in $\Phi$ generate a \defn{finite Coxeter group} $W$.
We denote by~$s_\alpha \in W$ the reflection orthogonal to a root~$\alpha \in \Phi$, and by~$\alpha_s \in \Phi^+$ the positive root orthogonal to a reflection~$s \in W$.
The group~$W$ is actually generated by the simple reflections~$S = \set{s_\alpha}{\alpha \in \Delta}$, and the pair~$(W,S)$ is a \defn{finite Coxeter system}.
Finite root systems and finite Coxeter systems are classified in terms of Dynkin diagrams, see~\cite{Humphreys}.

@@

It is known that~$\subwordComplex(Q,\omega)$ is either a ball or a sphere, in particular it is a pseudomanifold (with or without boundary).
The flip graph of~$\subwordComplex(Q,\omega)$ is the graph whose vertices are the facets of~$\subwordComplex(Q,\omega)$ and whose edges are the ridges of~$\subwordComplex(Q,\omega)$.
In other words, two facets $I,J \in \subwordFacets(Q,\omega)$ are connected by a \defn{flip} if they differ one element: $I \ssm \{i\} = J \ssm \{j\}$ for some $i\in I$ and $j\in J$ with~$i \ne j$.
The flip from $I$ to $J$ is called \defn{increasing} if~$i < j$, and \defn{decreasing} otherwise.
The \defn{increasing flip graph} on the facets of the subword complex is an acyclic graph which has a unique source and a unique sink~\cite{Pilaud-greedyFlipTree, PilaudStump-ELlabelings}.
These two special facets are called the \defn{greedy facet} $\greedyFacet$ and the \defn{antigreedy facet} $\antiGreedyFacet$.
They are the lexicographically smallest and largest facets of $\subwordComplex(Q,\pi)$.
The \defn{increasing flip poset} is the transitive closure of the increasing flip graph.

@@

\begin{example}
\label{exm:typeARootConfiguration}
Continuing \cref{exm:typeACoxeterSystem}, we consider the type~$A_n$ Coxeter system, the word $Q \eqdef \tau_1 \cdots \tau_{n-1} \tau_1 \cdots \tau_{n-2} \cdots \tau_1 \tau_2 \tau_1$, and a permutation~$\omega \in \fS_n$.
The word~$Q$ naturally fits on an $n \times n$ triangular grid (place $\tau_k$ in all boxes with row~$i$ and column~$j$ such that~$i+j+k = n+1$) and each facet~$I$ of the subword complex~$\subwordComplex(Q, \omega)$ corresponds to a pipe dream~$P_I$ of~$\pipeDreams(\omega)$ (replace each position in~$I$ by a contact in~$P_I$, and the other positions by crossings in~$P_I$).
The root function is given~$\rootFunction{I}{i} = \b{r}_{p,q} \eqdef \b{e}_p - \b{e}_q$ where~$p$ is the pipe arriving from the west and $q$ is the pipe arriving from the south at the box of~$P_I$ corresponding to position~$i$ of~$Q$.
Hence, the root configuration is the incidence configuration~$\Roots(I) = \bigset{\b{r}_{p,q}}{(p,q) \in P_I\contact}$ of the contact graph of~$P_I$.
In particular, the acyclic facets of~$\subwordComplex(Q, \omega)$ correspond to the acyclic pipe dreams of~$\acyclicPipeDreams(\omega)$.
\end{example}

@@

\begin{example}
\label{exm:counterExampleLattice}
In contrast to the previous example, consider the word~$Q = \tau_1 \tau_2 \tau_3 \tau_2 \tau_1 \tau_2 \tau_3 \tau_2 \tau_1$ on the simple generators of the symmetric group~$\fS_4$ and the subword complex~$\subwordComplex(Q,\wo)$.
Note that all facets in this subword complex are acyclic.
It is shown in~\cite[Rem.~5.12]{PilaudStump-brickPolytope} that its increasing flip poset is not a lattice.
This shows that neither the increasing flip poset, nor the acyclic increasing flip poset are lattices in general, even in type~$A$.
\end{example}

@@

\begin{proof}
%Let $\Phi = \Phi^+ \sqcup \Phi^-$ be the root system associated to $W$.
A classical result for finite root systems states that for any generic linear halfspace~$H^+$ the intersection~$\Phi \cap H^+$ is of the form $\pi(\Phi^+)$ for some $\pi \in W$.
Since any pointed cone is contained in some generic linear halfspace, we obtain
\[
\begin{array}[b]{rcl}
I \text{ is acyclic}  & \longleftrightarrow  &  \cone \Roots(I) \text{ is pointed} \\
& \longleftrightarrow  & \cone \Roots(I)\subseteq \pi(\Phi^+) \text{ for some } \pi\in W \\
& \longleftrightarrow  &  \Roots(I)\subseteq \pi(\Phi^+) \text{ for some } \pi\in W \\
& \longleftrightarrow  &  \linearExtensions(I) \neq \varnothing 
\end{array}
\qedhere
\]
\end{proof}

@@

A subset $\fR$ of the weak order is 
\begin{itemize}
\item a \defn{lower set} if $\sigma < \tau$ and $\tau \in \fR$ implies~$\sigma \in \fR$,
\item \defn{order convex} if $\sigma < \tau < \rho$ and~$\sigma, \rho \in \fR$ implies~$\tau \in \fR$.
\end{itemize}

@@

Now, since $\beta\notin \pi(\Phi^+)$.
There must be only one $\beta$ in the list.
Otherwise, there would be at least one $\beta$ whose position belongs to the facet $I$.
This would imply that $\beta\in \Roots(I)$ and $\pi$ would not be a linear extension of $I$, which is a contradiction.
So, our restricted list corresponding to $I$ looks like
\[
\begin{array}{ccc}
  \beta &, -\beta, \dots, &-\beta.\\
      i && j
\end{array}
\]
By  \cref{lem:rootFunctionFlips}~\eqref{lem:rootFunctionFlips3}, we obtain that flipping $j$ to $i$ creates the new facet $I' = I \ssm \{j\} \cap \{i\}$ whose corresponding restricted list looks like
\[
\begin{array}{ccc}
  \beta &, \beta, \dots, &\beta.\\
      i && j
\end{array}    
\]

@@

In this section, we present conjectural generalizations of the results of \cref{subsec:pipeDreamCongruence,subsec:pipeDreamQuotient}.
By \cref{coro:linearExtensionsPartitionSubwordComplexes}, we have
\[
\linearExtensions(Q,\omega) = \bigsqcup_{I \in \subwordAcyclicFacets(Q, \omega)} \linearExtensions(I)
\]
which naturally defines an equivalence relation on~$\linearExtensions(Q,\omega)$.
However, this equivalence relation is in general not a lattice congruence for two obvious reasons:
\begin{enumerate}[(i)]
\item while~$\linearExtensions(Q,\omega)$ is a lower set of the weak order containing~$[e, \omega]$ by \cref{thm:linearExtensionsPartitionSubwordComplexA}, it does not always coincides with~$[e, \omega]$ by \cref{rem:linearExtensionsPartitionSubwordComplexA}, and in fact it does not necessarily have a maximal element, hence it is not necessarily a lattice,
\item while the sets~$\linearExtensions(I)$ are always order convex in the weak order by \cref{thm:linearExtensionsPartitionSubwordComplexA}~\eqref{item:convex}, they are not necessarily intervals in the weak order by \cref{rem:linearExtensionsPartitionSubwordComplexA}.
\end{enumerate}
To bypass Issue~(i), we could restrict our attention to the situation when the word~$Q$ is sorting by~\cref{thm:linearExtensionsPartitionSubwordComplexB} (we will do that in \cref{conj:latticeQuotientSortingAlternating}).
Here, we want to be slightly more general, so we will instead consider general words~$Q$, but restrict our attention to the interval~$[e, \omega]$ as follows.
For a facet~$I$ of~$\subwordComplex(Q, \omega)$, we define~$\strongLinearExtensions(I) \eqdef \linearExtensions(I) \cap [e, \omega]$.
We say that~$I$ is \defn{strongly acyclic} if~$\strongLinearExtensions(I) \ne \varnothing$, and we denote by~$\subwordStronglyAcyclicFacets(Q, \omega)$ the set of strongly acyclic facets of~$\subwordComplex(Q, \omega)$.
Note that by \cref{thm:linearExtensionsPartitionSubwordComplexA}~\eqref{item:partition}, we have
\[
[e, \omega] = \bigsqcup_{I\in\subwordStronglyAcyclicFacets(Q,\omega)} \strongLinearExtensions(I).
\]
We can thus now define the analogue of the pipe dream congruence of \cref{def:pipeDreamCongruence} for subword complexes as follows.

@@

\subsection{The sweeping algorithm}
\label{subsec:sweepingAlgorithmSubwordComplexes}

We now extend the sweeping algorithm of \cref{subsec:sweepingAlgorithm} to construct, from a linear extension $\pi \in \linearExtensions(Q,\omega)$, the unique acyclic facet~$I$ such that $\pi \in \linearExtensions(I)$.
We note that this algorithm was independently described in~\cite{JahnStump}.

We start from a subword complex~$\subwordComplex(Q, \omega)$ and any element~$\pi \in W$.
The sweeping algorithm outcomes a facet for every element~$\pi \in W$, and this will be the desired facet for each $\pi \in \linearExtensions(Q,\omega)$.
In other words, we define a function
\[
\begin{tabular}{cccl}
$\sweepingAlgorithm$ : & $W$ & $\longrightarrow$ & $\subwordFacets(Q,\omega)$\\
& $\pi$ & $\longrightarrow$ & $\sweepingAlgorithm(Q,\omega,\pi)$
\end{tabular}
\]
\vincent{delete that}

It starts by setting $I^0=\varnothing$ to be the empty set.
It then scans the word $Q=(q_1,\dots ,q_m)$ from left to right.
At position $j\in [m]$ it produces a new set $I^j$ obtained from the previous one $I^{j-1}$ by either adding $j$ or not, according to certain rules.
Roughly speaking, the goal of the algorithm is to insert a reduced expression of $\omega$ in $Q$ by sweeping the word from left to right, while deciding whether we use a letter or not by looking at the root $\beta=\rootFunction{I^j}{j}$ that it produces.
If $\beta \in \Ninv(\omega)$ then it can not be used (Case~\ref{sweeping_case1}), otherwise we would not have a reduced expression for $\omega$.
If $\beta \in \Inv(\omega)$ then it takes it when $\beta \in \Inv(\pi)$ (Case~\ref{sweeping_case2}), or delays it as much as possible when $\beta \in \Ninv(\pi)$ (Case~\ref{sweeping_case3}).
If $\beta \in -\Inv(\omega)$ it can simply not take it (Case~\ref{sweeping_case4}).

More precisely, let $Q^j=Q_{[j]\ssm I^j}$ and denote by
\[
\rootFunction{I^j}{k} = \prod Q_{[k-1]\ssm I^j}(\alpha_{q_k}).
\]
the partial root function for $k\leq j$.  
The rules are the following:

\begin{enumerate}[{Case} 1:]
    \item If $\rootFunction{I^j}{j}\in \Ninv(\omega)$ then 
    \(
    I^j=I^{j-1}\cup\{j\}.
    \) 
    \label{sweeping_case1}
    \item If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Inv(\pi)$ then 
    \(
    I^j=I^{j-1}.
    \)
    \label{sweeping_case2}
    \item If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Ninv(\pi)$ then we consider two cases: 
    \label{sweeping_case3} 
    \begin{enumerate}[\hspace{-1cm} (a)]
        \item If $Q_{[m]\ssm I^{j-1}\ssm \{j\}}$ contains a reduced expression of $\omega$ with prefix $Q^{j-1}$ then 
        \(
        {I^j=I^{j-1}\cup \{j\}}.
        \)~
        \label{sweeping_case3a}
        \item Otherwise
        \(
        I^j=I^{j-1}.
        \)
        \label{sweeping_case3b}
    \end{enumerate}
    \item If $\rootFunction{I^j}{j}\in -\Inv(\omega)$ then
    \(
    I^j=I^{j-1}\cup \{j\}.
    \)
    \label{sweeping_case4}
\end{enumerate}

We denote by $\sweepingAlgorithm(Q,\omega,\pi)$ the resulting set $I^m$ obtained at the last step of the algorithm.
The objective of this section is the following statement.

\begin{proposition}
\label{prop:sweeping}
Let $\subwordComplex(Q,\omega)$ be a non-empty subword complex. Then, for every $\pi \in W$, 
\begin{enumerate}
\item the set $\sweepingAlgorithm(Q,\omega,\pi)$ is a facet of $\subwordComplex(Q,\omega)$, \label{item:sweeping1}
\item if $\pi\in \linearExtensions(I)$ for some facet $I\in \subwordComplex(Q,\omega)$, then $\sweepingAlgorithm(Q,\omega,\pi) = I$. \label{item:sweeping2}
\end{enumerate}
\end{proposition}

We first need to argue that the algorithm terminates, that is, that any position $j$ falls into one of the four cases above.

\begin{lemma}
\label{lem:sweeping1}
At step $j$ of the sweeping algorithm, the root~$\rootFunction{I^j}{j}$ belongs to exactly one of~the~sets
\[
\Ninv(\omega),
\qquad
\Inv(\omega) \cap \Inv(\pi),
\qquad
\Inv(\omega) \cap \Ninv(\pi)
\qquad\text{or}\qquad
-\Inv(\omega).
\]
\end{lemma}

\begin{proof}
The root $\rootFunction{I^j}{j}$ belongs to~$\Phi^+ \sqcup \Phi^-$ and we know that
\[
\Phi^+ = \Inv(\omega) \sqcup \Ninv(\omega) \qquad \qquad
\Phi^- = -\Inv(\omega) \sqcup -\Ninv(\omega)  
\]
%The positive roots are covered by the cases (1),(2), and (3).
%The negative roots in $-\Inv(\omega)$ are covered in case (4).
Since we are inserting a reduced expression of $\omega$ in $Q$, the case $-\Ninv(\omega)$ never occurs.
\end{proof}

We now observe the main invariant of the sweeping algorithm.

\begin{lemma}
\label{lem:sweeping2}
At any step~$j$ of the algorithm, $Q^j$ is a reduced expression which is the prefix of a reduced expression of $\omega$ in $Q_{[m]\ssm I^j}$.
\end{lemma}

\begin{proof}
The proof works by induction on $j$.

@@

We assume now that the claim holds for $j-1$ and we prove it for $j$.
Note that 
\[
Q^j = 
\begin{cases}
Q^{j-1}, & \text{if } I^j=I^{j-1}\cup \{j\} ,\\
Q^{j-1} q_j, & \text{if } I^j=I^{j-1}.
\end{cases}
\]
We analyze the different cases of the sweeping algorithm.
\begin{enumerate}
    \item If $\rootFunction{I^j}{j}\in \Ninv(\omega)$ then 
    $Q^j=Q^{j-1}$, which is reduced by induction hypothesis.
    Moreover, it is a prefix of a reduced expression of $\omega$ in $Q_{[m]\ssm I^{j-1}}$.
    Since $\rootFunction{I^j}{j}\in \Ninv(\omega)$, this reduced expression can not use the letter $q_j$; so, it is a reduced expression of $\omega$ in $Q_{[m]\ssm I^{j}}$ as desired.
    \item If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Inv(\pi)$ then $Q^j= Q^{j-1}\circ (q_j)$, which is a reduced expression because~$Q^{j-1}$ is reduced and $\rootFunction{I^j}{j}\in \Inv(\omega)$.

@@

    \item[(3)(a)] If $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Ninv(\pi)$ and $Q_{[m]\ssm I^{j-1}\ssm \{j\}}$ contains a reduced expression of $\omega$ with prefix $Q^{j-1}$, then $Q^j=Q^{j-1}$.
    Therefore, $Q^j$ is reduced and it is a prefix of a reduced expression of $\omega$ in 
    $Q_{[m]\ssm I^{j-1}\ssm \{j\}}=Q_{[m]\ssm I^j}$.

    \item[(3)(b)] In this case, $\rootFunction{I^j}{j}\in \Inv(\omega) \cap \Ninv(\pi)$ but $Q^j= Q^{j-1}\circ (q_j)$. The proof is similar to the proof of (2).

    \item[(4)] If $\rootFunction{I^j}{j}\in -\Inv(\omega)$ then $Q^j=Q^{j-1}$. The argument is similar to that of Case~(1).
    \qedhere
    \end{enumerate}
\end{proof}

Finally, we need the following technical statement.

\begin{lemma}
\label{lem:sweeping3}
Let $I_1,I_2\in \subwordComplex(Q,\omega)$ be two different facets, and $j\in [m]$ be the first position where they differ.
Without loss of generality assume 
\[
I_2\cap [j] = I_1\cap [j] \ssm \{j\}
\]
with $j\in I_1$.
Let $\beta=\rootFunction{I_1}{j}=\rootFunction{I_2}{j}$, then 
\[
-\beta \in \cone \Roots(I_2).
\]
\end{lemma}

@@

\begin{proof}[Proof of \cref{prop:sweeping}]
Since the sweeping algorithm terminates by \cref{lem:sweeping1}, Point~\eqref{item:sweeping1} follows directly from the invariant of the sweeping algorithm of \cref{lem:sweeping2} applied when~$j = m$.

For Point~\eqref{item:sweeping2}, assume $\pi \in \linearExtensions(I)$ and let $I^j=I\cap [j]$.
We will show that the partial root function~$\rootFunction{I^j}{\cdot}$ agrees with the decisions taken in the sweeping algorithm.
Indeed, we will see that those decisions are forced.
